---
title: "Lab 8"
author: "Marin Azhar"
output:
  pdf_document: default
  latex_engine: xelatex
date: "11:59PM April 29, 2021"
---
  
  I want to make some use of my CART package. Everyone please try to run the following:
  
```{r}
#if (!pacman::p_isinstalled(YARF)){
#  pacman::p_install_gh("kapelner/YARF/YARFJARs", ref = "dev")
 # pacman::p_install_gh("kapelner/YARF/YARF", ref = "dev", force = TRUE)
#}
#options(java.parameters = "-Xmx4000m")
#pacman::p_load(YARF)
```

For many of you it will not work. That's okay.

Throughout this part of this assignment you can use either the `tidyverse` package suite or `data.table` to answer but not base R. You can mix `data.table` with `magrittr` piping if you wish but don't go back and forth between `tbl_df`'s and `data.table` objects.

```{r}

pacman::p_load(tidyverse, magrittr, data.table)
```

We will be using the `storms` dataset from the `dplyr` package. Filter this dataset on all storms that have no missing measurements for the two diameter variables, "ts_diameter" and "hu_diameter".

```{r}
#TO-DO
data(storms)
storms2 = storms%>%
  filter(!is.na(ts_diameter )& !is.na(hu_diameter ) & ts_diameter> 0 & hu_diameter >0)

storms2
```

From this subset, create a data frame that only has storm, observation period number for each storm (i.e., 1, 2, ..., T) and the "ts_diameter" and "hu_diameter" metrics.

```{r}
#TO-DO

storms2 = storms2%>% 
  select(name, ts_diameter, hu_diameter)%>%
  group_by(name)%>%
  mutate(period = row_number())
  
```

Create a data frame in long format with columns "diameter" for the measurement and "diameter_type" which will be categorical taking on the values "hu" or "ts".

```{r}
#TO-DO
storms_long =pivot_longer(storms2, cols =matches("diameter"), names_to = "diameter")
storms_long 
#the cols argument tells the pivot longer which of the cols to consolidate to one long col

```

Using this long-formatted data frame, use a line plot to illustrate both "ts_diameter" and "hu_diameter" metrics by observation period for four random storms using a 2x2 faceting. The two diameters should appear in two different colors and there should be an appropriate legend.

```{r}
#TO-DO
storms_sample = sample(unique(storms2$name), 4)

ggplot(storms_long%>% filter(name %in% storms_sample))+
  geom_line(aes(x=period, y=value, col = diameter))+          #we squished all four storms....
  facet_wrap(name ~., nrow =2)

```

In this next first part of this lab, we will be joining three datasets in an effort to make a design matrix that predicts if a bill will be paid on time. Clean up and load up the three files. Then I'll rename a few features and then we can examine the data frames:
  
```{r}
rm(list = ls())
pacman::p_load(tidyverse, magrittr, data.table, R.utils)
bills = fread("https://github.com/kapelner/QC_MATH_342W_Spring_2021/raw/master/labs/bills_dataset/bills.csv.bz2")
payments = fread("https://github.com/kapelner/QC_MATH_342W_Spring_2021/raw/master/labs/bills_dataset/payments.csv.bz2")
discounts = fread("https://github.com/kapelner/QC_MATH_342W_Spring_2021/raw/master/labs/bills_dataset/discounts.csv.bz2")
setnames(bills, "amount", "tot_amount")
setnames(payments, "amount", "paid_amount")
head(bills)
head(payments)
head(discounts)
#the data is fake data
#bill data and payment data has a relation 

bills =as_tibble(bills) #data.table to 
payments =as_tibble(payments)
discounts =as_tibble(discounts)
```

The unit we care about is the bill. The y metric we care about will be "paid in full" which is 1 if the company paid their total amount (we will generate this y metric later).

Since this is the response, we would like to construct the very best design matrix in order to predict y.

I will create the basic steps for you guys. First, join the three datasets in an intelligent way. You will need to examine the datasets beforehand.

```{r}
bills_with_payments = left_join(bills, payments, by = c("id" = "bill_id"), all.x = TRUE)
#id payments = id.y
bills_with_payments
  bills_with_payments_with_discounts = left_join(bills_with_payments, discounts, by = c("discount_id" ="id" ), all.x = TRUE)
```

Now create the binary response metric `paid_in_full` as the last column and create the beginnings of a design matrix `bills_data`. Ensure the unit / observation is bill i.e. each row should be one bill! 
  
```{r}
#TO-DO
bills_data = bills_with_payments_with_discounts%>%
  mutate(tot_amount = if_else(is.na(pct_off), tot_amount, tot_amount*(1-pct_off/100)))%>%
  group_by(id)%>%
  mutate(sum_of_payment_amount = sum(paid_amount))%>%
  mutate(paid_in_full = if_else(sum_of_payment_amount >= tot_amount, 1,0, missing =0  ))%>%
  slice(1)%>%
  ungroup()
table(bills_data$paid_in_full, useNA = "always") 
```

How should you add features from transformations (called "featurization")? What data type(s) should they be? Make some features below if you think of any useful ones. Name the columns appropriately so another data scientist can easily understand what information is in your variables.

```{r}
#TO-DO

  
  pacman::p_load("lubridate")
  bills_data = bills_data %>%
  select(-id, -id.y, -num_days, -transaction_date, -pct_off, -days_until_discount, -sum_of_payment_amount, -paid_amount) %>%
  mutate(num_days_to_pay = as.integer(ymd(due_date) - ymd(invoice_date))) %>%
  select(-due_date, -invoice_date) %>%
  mutate(discount_id = as.factor(discount_id)) %>%
  group_by(customer_id) %>%
  mutate(bill_num = row_number()) %>%
  ungroup() %>%
  select(-customer_id) %>%
  relocate(paid_in_full, .after = last_col())
```

Now let's do this exercise. Let's retain 25% of our data for test.

```{r}
K = 4
test_indices = sample(1 : nrow(bills_data), round(nrow(bills_data) / K))
train_indices = setdiff(1 : nrow(bills_data), test_indices)
bills_data_test = bills_data[test_indices, ]
bills_data_train = bills_data[train_indices, ]
```

Now try to build a classification tree model for `paid_in_full` with the features (use the `Xy` parameter in `YARF`). If you cannot get `YARF` to install, use the package `rpart` (the standard R tree package) instead. You will need to install it and read through some documentation to find the correct syntax.

Warning: this data is highly anonymized and there is likely zero signal! So don't expect to get predictive accuracy. The value of the exercise is in the practice. I think this exercise (with the joining exercise above) may be one of the most useful exercises in the entire semester.

```{r}
#TO-DO
pacman::p_load(rpart)
mod1 = rpart(paid_in_full ~., data = bills_data_train, method = "class")
```

For those of you who installed `YARF`, what are the number of nodes and depth of the tree? 

```{r}
#TO-DO
nrow(mod1$frame) ##number of nodes
```

For those of you who installed `YARF`, print out an image of the tree.

```{r}
#TO-DO
plot(mod1, uniform=TRUE)
text(mod1, use.n=TRUE, all=TRUE, cex=.8)
```

Predict on the test set and compute a confusion matrix.

```{r}
#TO-DO
yhat = predict(mod1, bills_data_test, type = c("class"), na.action = na.pass)
oos_conf_table=table( bills_data_test$paid_in_full,yhat)
#yhat is top most row

```

Report the following error metrics: misclassifcation error, precision, recall, F1, FDR, FOR.

```{r}
#TO-DO
n = sum(oos_conf_table)
fp = oos_conf_table[1, 2]
fn = oos_conf_table[2, 1]
tp = oos_conf_table[2, 2]
tn = oos_conf_table[1, 1]
num_pred_pos = sum(oos_conf_table[, 2])
num_pred_neg = sum(oos_conf_table[, 1])
num_pos = sum(oos_conf_table[2, ])
num_neg = sum(oos_conf_table[1, ])
me= (fn+fp)/n
cat("misclassifcation error", round(me*100,2), "%\n")
precision = tp / num_pred_pos
cat("precision", round(precision * 100, 2), "%\n")
recall = tp / num_pos
cat("recall", round(recall * 100, 2), "%\n") #true positive in relation to false negatives 
false_discovery_rate = 1 - precision
cat("false_discovery_rate", round(false_discovery_rate * 100, 2), "%\n") #false pos???????
false_omission_rate = fn / num_pred_neg
cat("false_omission_rate", round(false_omission_rate * 100, 2), "%\n") # false negative 
#lec 21
```

Is this a good model? (yes/no and explain).
#true positives over all postives 
#TO-DO
no, it is not a good model the error rates are higher than what we want
because the fdr is 1/3 which is bad it means 2/3 of the ppl are not paying 

There are probability asymmetric costs to the two types of errors. Assign the costs below and calculate oos total cost.

```{r}
#TO-DO
c_fp =45
c_fn =2
cost =c_fp*fp + c_fn*fn
cost
```

We now wish to do asymmetric cost classification. Fit a logistic regression model to this data.

```{r}

log_mod =glm(paid_in_full~., bills_data_train, family = binomial(link ="logit"))
#p_hat_train = predict(log_mod, bills_data_train, type ="response")

```

Use the function from class to calculate all the error metrics for the values of the probability threshold being 0.001, 0.002, ..., 0.999 in a data frame.

```{r}

compute_metrics_prob_classifier = function(p_hats, y_true, res = 0.001){
  #we first make the grid of all prob thresholds
  p_thresholds = seq(0 + res, 1 - res, by = res) #values of 0 or 1 are trivial
  
  #now we create a matrix which will house all of our results
  performance_metrics = matrix(NA, nrow = length(p_thresholds), ncol = 12)
  colnames(performance_metrics) = c(
    "p_th",
    "TN",
    "FP",
    "FN",
    "TP",
    "miscl_err",
    "precision",
    "recall",
    "FDR",
    "FPR",
    "FOR",
    "miss_rate"
  )
  
  #now we iterate through each p_th and calculate all metrics about the classifier and save
  n = length(y_true)
  for (i in 1 : length(p_thresholds)){
    p_th = p_thresholds[i]
    y_hats = factor(ifelse(p_hats >= p_th, 1, 0))
    confusion_table = table(
      factor(y_true, levels = c(0, 1)),
      factor(y_hats, levels = c(0, 1))
    )
      
    fp = confusion_table[1, 2]
    fn = confusion_table[2, 1]
    tp = confusion_table[2, 2]
    tn = confusion_table[1, 1]
    npp = sum(confusion_table[, 2])
    npn = sum(confusion_table[, 1])
    np = sum(confusion_table[2, ])
    nn = sum(confusion_table[1, ])
  
    performance_metrics[i, ] = c(
      p_th,
      tn,
      fp,
      fn,
      tp,
      (fp + fn) / n,
      tp / npp, #precision
      tp / np,  #recall
      fp / npp, #false discovery rate (FDR)
      fp / nn,  #false positive rate (FPR)
      fn / npn, #false omission rate (FOR)
      fn / np   #miss rate
    )
  }
  
  #finally return the matrix
  performance_metrics
} 
  
p_hat_train = predict(log_mod, bills_data_train, type ="response")



p_hat_test = predict(log_mod, bills_data_test, type = "response")

performance_metrics_in_sample = as_tibble(compute_metrics_prob_classifier(p_hat_train,  bills_data_train$paid_in_full))
performance_metrics_in_sample

performance_metrics_oos = as_tibble(compute_metrics_prob_classifier(p_hat_test, bills_data_test$paid_in_full)) 
performance_metrics_oos

#metrics_prob_classifier = compute_metrics_prob_classifier(p_hat_train, y_true)
#metrics_prob_classifier2=as_tibble(metrics_prob_classifier)
#metrics_prob_classifier2

```

Calculate the column `total_cost` and append it to this data frame.

```{r}
c_fp =45
c_fn =2
#metrics_prob_classifier2= metrics_prob_classifier2%>%
performance_metrics_oos=performance_metrics_oos%>% 
 mutate(total_cost =c_fp*FP + c_fn*FN)
performance_metrics_oos

performance_metrics_in_sample=performance_metrics_in_sample%>% 
 mutate(total_cost =c_fp*FP + c_fn*FN)
performance_metrics_in_sample
#metrics_prob_classifier2
#TO-DO
```

Which is the winning probability threshold value and the total cost at that threshold?

```{r}
#TO-DO
best_prob_thsehold_index2 = which.min(performance_metrics_oos$total_cost)
best_prob_thsehold_index2
best_prob_thsehold_matrix2 = performance_metrics_oos[best_prob_thsehold_index2,]
best_prob_thsehold_matrix2

cat("total cost of win probs oos", min(best_prob_thsehold_matrix2$total_cost))


```

Plot an ROC curve and interpret.

```{r}
#TO-DO

pacman::p_load(ggplot2)
#ggplot(performance_metrics_in_sample) +
#  geom_line(aes(x = FPR, y = recall)) +
# geom_abline(intercept = 0, slope = 1, col = "orange") + 
# coord_fixed() + xlim(0, 1) + ylim(0, 1)

#now do for the oos + in sample

performance_metrics_in_and_oos= rbind(
    cbind(performance_metrics_in_sample, tibble(sample = "in")),
   cbind(performance_metrics_oos, tibble(sample = "out"))
)




ggplot(performance_metrics_in_and_oos) +
  geom_line(aes(x = FPR, y = recall, col = sample)) +
  geom_abline(intercept = 0, slope = 1, col = "orange") + 
  coord_fixed() + xlim(0, 1) + ylim(0, 1)


#out of sample has higher recall compared to in sample but there is a trade off because the insample has higher fpr while the out of sample has lower fpr 
```

#TO-DO interpretation

Calculate AUC and interpret.

```{r}
#TO-DO

pacman::p_load(pracma)
-trapz(performance_metrics_in_and_oos$FPR, performance_metrics_in_and_oos$recall)

#the AUC is closer to 1 which means the model is able distinguish/ predict when a person will pay or not pay their bills. Thus this is a good model.
```

#TO-DO interpretation

Plot a DET curve and interpret.

```{r}
#TO-DO
table(bills_data_test$paid_in_full) / length(bills_data_test$paid_in_full)
table(bills_data_train$paid_in_full) / length(bills_data_train$paid_in_full)




#this is funky

ggplot(performance_metrics_in_and_oos) +
  geom_line(aes(x = FDR, y = miss_rate, col = sample)) +
  coord_fixed() + xlim(0, 1) + ylim(0, 1)
```

#TO-DO interpretation


Here we are calculation for the error rates for the binary classification for false positives and flase negatives 
it's suppose to look weird but it's helpful.


