---
title: "Final Project.Rmd"
author: "Marin Azhar"
output:
  word_document: default
  html_document: default
---
#Loading the 2016-2017 Housing Data for Queens
```{r}
pacman::p_load(tidyverse, magrittr, data.table, R.utils, skimr)
housing_data = fread("https://raw.githubusercontent.com/kapelner/QC_MATH_342W_Spring_2021/master/writing_assignments/housing_data_2016_2017.csv")

housing_data = data.frame(housing_data)

```

We have 55 cols and 2230 rows in total.
Our repose variable is sales_price.

#Now we are dropping data that is not relevant to our data set.

```{r}
housing_data = housing_data%<>%
  select(-HITId, -HITTypeId, -Title, -Description, -Keywords, -Reward, -CreationTime, -MaxAssignments, -RequesterAnnotation, -AssignmentDurationInSeconds, -AutoApprovalDelayInSeconds, -NumberOfSimilarHITs, -LifetimeInSeconds, -AssignmentId, -WorkerId, -AssignmentStatus, -AcceptTime, -SubmitTime, -AutoApprovalTime, -ApprovalTime,  -RejectionTime, -RequesterFeedback, -URL, -url, -Expiration, -Last30DaysApprovalRate, -Last7DaysApprovalRate ,-WorkTimeInSeconds, -LifetimeApprovalRate )

#colnames(housing_data)
#skim(housing_data)
```


#After another closer look at the data:

1) We can drop model type because there are 875 unique categories and the information is either redundant (about number of bedrooms and bathrooms) or not useful because it's about a very specific building type name which has no value to us because we do not have the time to combined categories or gain any extra information from this col.  

2) We can drop listing_price_to_nearest_1000 because it's redundant information since we have the sales_price col.

3) We can drop num_floors_in_building because it is unnecessary and nyc has a building regulation (Section BC 3002.4) any building with more than five stories must have an elevator. While the data set does have buildings with less than five floors but because 650 rows have missing data and we can not impute or guess the number of floors a building has we will just drop this feature all together.

etc.

```{r}
#after looking out our data set there are extra cols that we can delete 
housing_data = housing_data%<>%
  select( -model_type , -listing_price_to_nearest_1000, -num_floors_in_building, -pct_tax_deductibl, -date_of_sale, -community_district_num  )

#-fuel_type
```

#Cleaning the data to reduce unique catigories 

```{r}
#notes taken with Sara on friday
#find out deal with LifetimeApprovalRate
#make approx_year_built into categorical ranges
#binarize cats_allowed/dogs_allowed into 0's and 1's
#make community district number into categorical features (factor)
#make coop and condo into 0's and 1's
#dining room type into categorical with reference category
#combine half bathrooms with full bathrooms into one feature (maybe not?)
#get rid of model_type col, but potentially use the values it contains to fill in blanks missing for other features (?)
#combine the two types of "other" categories for fuel_type
#take out either the zip of the address of full_address_or_zip_code
#fix garage_exists and combine yes, Yes, eys, 1, UG, Underground into a single category 
#fix kitchen type - combine all of them into "efficiency", "combo", "eatin", "none" (get rid of 1955?!?)
#take out the dollar sign in maintnence_cost and make it into a numerical/continuous feature (it's currently a character)
#num_total_rooms - is it worth keeping if we have num_bedrooms, number of bathrooms and whether or not there's a kitchen and dining room?
#sale_price take out dollar signs and convert to a numerical/int type
#sq_footage - should we make continuous or categorical range of values - make sure to impute the missingness (with missForest)
#listing_price_to_nearest_1000 needs to have $ removed, and converted from chr to int/numerical

#colnames(housing_data)
```

#Create zipcodes to catigorize the locations 

```{r}
#from Janine
zip_codes = gsub("[^0-9.-]", "", housing_data$full_address_or_zip_code)
housing_data$zip_codes = str_sub(zip_codes, -5, -1)

```

#Clean Zipcodes

```{r}
#unique(housing_data$zip_codes) # check for any incorrect zipcodes 
 housing_data%>%
  filter( housing_data$zip_codes =="1355." |  housing_data$zip_codes == "1367." | housing_data$zip_codes == "17-30" | housing_data$zip_codes == ".1136" | housing_data$zip_codes =="71137" |  housing_data$zip_codes == "01137" | housing_data$zip_codes == "81137" | housing_data$zip_codes == "01137" | housing_data$zip_codes == "71136" | housing_data$zip_codes == "51142"| housing_data$zip_codes == "51135")%>% #find correct zips and manually change them 
  select(full_address_or_zip_code, zip_codes)

#technically we could delete them since it's only 7 rows but because it's only 7 rows i can also manually fix it
 


#Clean Zipcodes



housing_data$zip_codes[housing_data$zip_codes == "1367."] <-"11367"
#housing_data$zip_codes[housing_data$zip_codes == "17-30"] <-NA no address 
housing_data$zip_codes[housing_data$zip_codes == ".1136"] <-"11369"
housing_data$zip_codes[housing_data$zip_codes == "1355."] <-"11355"

housing_data$zip_codes[housing_data$zip_codes == "81137"] <-"11372" 
housing_data$zip_codes[housing_data$zip_codes == "71137"] <-"11372"
housing_data$zip_codes[housing_data$zip_codes == "01137"] <-"11375"
housing_data$zip_codes[housing_data$zip_codes == "1136"] <-"11364"
housing_data$zip_codes[housing_data$zip_codes == "51142"] <-"11427"

housing_data$zip_codes[housing_data$zip_codes == "51135"] <-"11355"
housing_data$zip_codes[housing_data$zip_codes == "71136"] <-"11364"


  
housing_data = housing_data[housing_data$zip_codes !="17-30",] #remove rows
#unique(housing_data$zip_codes)
```


```{r}
#Mutate Zipcodes into Neigbhorhoods


house2 = housing_data %>%
  mutate(
    
 zip_codes =    ifelse(zip_codes == "11361" | zip_codes == "11362" | zip_codes == "11363" | zip_codes == "11364" , "Northeast Queens",zip_codes), 
  zip_codes =  ifelse( zip_codes == "11354" | zip_codes == "11355" | zip_codes == "11356" | zip_codes == "11357" | zip_codes == "11358" | zip_codes == "11359" | zip_codes == "11360" , "North Queens", zip_codes),
         
 zip_codes =   ifelse( zip_codes == "11365" | zip_codes == "11366" | zip_codes == "11367" ,"Central Queens",zip_codes),
   
zip_codes =    ifelse( zip_codes == "11412" | zip_codes == "11423" | zip_codes == "11432" | zip_codes == "11433" | zip_codes == "11434" | zip_codes == "11435" | zip_codes == "11436" , "Jamaica" , zip_codes) ,
   
zip_codes =    ifelse( zip_codes == "11101" | zip_codes == "11102" | zip_codes == "11103" | zip_codes == "11104" | zip_codes == "11105" | zip_codes == "11106","Northwest Queens", zip_codes),
   
zip_codes =     ifelse(zip_codes == "11374" | zip_codes == "11375" | zip_codes == "11379" | zip_codes == "11385" , "West Central Queens", zip_codes),
   
 zip_codes =   ifelse(zip_codes == "11004" | zip_codes == "11005" | zip_codes == "11411" | zip_codes == "11413" | zip_codes == "11422" | zip_codes == "11426" | zip_codes == "11427" | zip_codes == "11428" | zip_codes == "11429" , "Southeast Queens", zip_codes),
   
 zip_codes =   ifelse(zip_codes == "11414" | zip_codes == "11415" | zip_codes == "11416" | zip_codes == "11417" | zip_codes == "11418" | zip_codes == "11419" | zip_codes == "11420"  | zip_codes == "11421" , "Southwest Queens", zip_codes),
    
zip_codes =    ifelse(zip_codes == "11368" | zip_codes == "11369" | zip_codes == "11370" | zip_codes == "11372" | zip_codes == "11373" | zip_codes == "11377" | zip_codes == "11378"  ,"West Queens", zip_codes)
    
    
    
)
#unique(house2$zip_codes)


```

#Mutate other features 

```{r}
house2%<>%
  mutate(
    
    cats_allowed = as.factor(ifelse(cats_allowed == "yes" | cats_allowed == "y", "yes","no")),
    
    coop_condo =as.factor(coop_condo),
    zip_codes = as.factor(zip_codes),
    
    dogs_allowed = as.factor(ifelse(dogs_allowed == "yes" | dogs_allowed == "yes89", "yes","no")),
    
    kitchen_type = as.factor(case_when(kitchen_type == "efficiency kitchen" | kitchen_type == "efficiency" | 
                            kitchen_type == "efficiemcy"| kitchen_type ==  "efficiency ktchen"| 
                            kitchen_type == "efficiency kitchene" ~"efficiency kitchen", 
                            kitchen_type == "eat in" | kitchen_type == "Eat In" | kitchen_type == "eatin" |  
                            kitchen_type =="1955"|kitchen_type == "Eat in" ~ "eat in", 
                            kitchen_type =="Combo"| kitchen_type == "combo" ~ "combo")),
    
    dining_room_type= as.factor(case_when(dining_room_type == "none" |
                                dining_room_type == "other" ~"other" , dining_room_type == "dining area" | 
                                dining_room_type =="combo" ~ "combo", dining_room_type =="formal" ~"formal")),
    
    fuel_type = as.factor(ifelse(fuel_type == "other" | fuel_type == "Other" | fuel_type =="none", "other", fuel_type)),        #not 
#using fuel type because the zipcode/location dictates what kind of infrastructure a building can built, there fore fuel type is a Spurious relationship....
   
    
    #change garage to a parking space a parking space that is non underground could exist and we can tell based on if a parking charges exist also how do we know if yes = underground or if it's different,thus make it all the same category 
    
    garage_exists = as.factor(ifelse(garage_exists =="Underground"| garage_exists == "Yes" | garage_exists == "yes" |
                              garage_exists == "1" |garage_exists =="eys"  | is.na(parking_charges) ==FALSE ,"yes","no")), #dealing with missingnes minimally 
  
    walk_score = as.factor(case_when( walk_score < 50 ~ "Car-Dependent", 
                                  walk_score >49 &walk_score <70 ~ "Somewhat Walkable",
                                  walk_score>69 & walk_score <90 ~ "Very Walkable", 
                                  walk_score>89 & walk_score <= 100 ~ "Walker's Paradise")), 
  

   # approx_year_built = as.factor(case_when(approx_year_built < 1939 ~ "Pre war", 
                                        #  approx_year_built >=1939  &  approx_year_built  <= 1990~ "During or Post war" ,
                                         # approx_year_built >=1990 ~ "Contemporary (after 1990)" )),
      
    common_charges = as.integer(str_remove_all(common_charges, "[$,]")),
    maintenance_cost = as.integer(str_remove_all(maintenance_cost, "[$,]")),
    total_taxes = as.integer(str_remove_all(total_taxes, "[$,]" )), # need na 
    total_taxes =(total_taxes/12), # make it a monthly cost
    sale_price = as.integer(str_remove_all(sale_price, "[$,]")) # this is our reponse varable 
    
   
   )

```


```{r}
 house2%<>%
select(-parking_charges, -full_address_or_zip_code )

```


```{r}
house2 %<>% 
  rename(
    parking_exists = garage_exists, #it could have a parking lot isntead of a garage moreover because we had the parking prices we know a spot somowhere existed 
    neighborhood = zip_codes
    )

house2$parking_exists[is.na(house2$parking_exists) ==TRUE ] <-"no" #we are making an assumption here
house2$num_half_bathrooms[is.na(house2$num_half_bathrooms) ==TRUE] <-0 # asumption

```


```{r}
#Problem sometimes the co_op's do not have a maintenance cost by they have a common charage, and vis versa one way to deal with this is to swap the charges in the right place. co_ops should only have a monthy maintance cost and condos should only have a monthy common chrage + total taxtes.


house2$fix_mc_swaps = rowSums(house2[ , c("common_charges","maintenance_cost")], na.rm =TRUE) 
house2$fix_cc_swaps_add_total_taxes = rowSums(house2[ ,c("common_charges","maintenance_cost", "total_taxes")], na.rm =TRUE) #maybe dont add total taxes yet 
house2$maintenance_cost = house2$fix_mc_swaps
house2$common_charges = house2$fix_cc_swaps_add_total_taxes
```


```{r}
#Fill back the Zeros


house_condo = house2%>%
filter(coop_condo =="condo")

house_coop =house2%>%
  filter(coop_condo == "co-op")

house_condo$maintenance_cost <- 0 #condos do not have maintenance_cost
house_condo$common_charges[house_condo$common_charge == 0] <- NA #to impute on common_charfes
house_coop$common_charges <- 0 #coops do not have common_charges
house_coop$maintenance_cost[house_coop$maintenance_cost == 0]<- NA #if a zero impute cost
```


```{r}
#Combind back the Data set


house2 =rbind(house_condo, house_coop) #stich back the two data frames 

house2 =house2%>%
  select(-fix_mc_swaps, -fix_cc_swaps_add_total_taxes, -total_taxes) # drop unecessary cols

house2 <- house2[sample(1:nrow(house2)), ] # to randmize the order of the data again 

```


```{r}
# final clean drop rows so features have 0% missingness 


house2 = house2[is.na(house2$common_charges) == FALSE,] #only one missing just drop row
#house2 = house2[is.na(house2$num_total_rooms) == FALSE,] #only two missing just drop row //dropped it so the lin model can work

summary(house2) #for part 2 in the essay 
```

#Create dummy varaiables for imputation on features that have less than 90% missing 

```{r}
set.seed(718)

M = tbl_df(apply(is.na(house2), 2, as.numeric))
colnames(M) = paste(colnames(house2), "_is_missing", sep = "")
M = tbl_df(t(unique(t(M))))
M %<>% 
  select_if(function(x){sum(x) > 0})

house2 = cbind(M, house2)
```
```{r}
missing_col = ncol(M)
```
#final list of features 
```{r}
colnames(house2)
```


#create train and test splits

```{r}
obs_without_responses = house2 %>%
  filter(is.na(sale_price))
obs_with_responses = house2 %>%
  filter(!is.na(sale_price))

n = nrow(obs_with_responses) #there are 528 observations with responses that we can use later on
k = 5

test_indices = sample(1 : n, 1 / k * n)
train_indices = setdiff(1 : n, test_indices)

n_test = as.integer((1 / k) * n)
n_train = as.integer(n - n_test)

training = obs_with_responses[train_indices, ]
testing = obs_with_responses[test_indices, ]

X_test = testing %>%
  mutate(sale_price = NA)
y_test = testing$sale_price

```

#impute on data using missForest


```{r}
pacman::p_load(missForest)

#fill in missingness
housing_missing = rbind(training, X_test, obs_without_responses) #can use all data except y_test (to use it would be cheating)

housing_complete = missForest(housing_missing)$ximp
#housing_complete
sum(is.na(housing_complete))
skim(housing_complete)

housing = housing_complete %>%
  filter(sale_price_is_missing == 0) %>%
  select(-sale_price_is_missing) 

housing = cbind(housing[, -(1:missing_col)], tbl_df(t(unique(t(housing[,(1:missing_col)]))))) #make sure all col are linearly independent

housing_training = housing[1:n_train, ]
housing_test = housing[(n_train+1):n, ]

housing_test$sale_price = y_test

```

```{r}
#combine charges with maintenance cost after imputation before creating models 
housing_test %<>%
  mutate(total_cost = maintenance_cost + common_charges) %<>%
  select(-maintenance_cost, -common_charges)

housing_training %<>%
  mutate(total_cost = maintenance_cost + common_charges) %<>%
  select(-maintenance_cost, -common_charges)

housing_ytest = housing_test$sale_price
housing_Xtest = housing_test
housing_Xtest$sale_price = NULL

housing_ytrain = housing_training$sale_price
housing_Xtrain = housing_training
housing_Xtrain$sale_price = NULL
```

```{r}
#Regression Tree Model
pacman::p_load(YARF)
options(java.parameters = "-Xmx4000m")

reg_tree = YARFCART(housing_Xtrain, housing_ytrain)
reg_tree
get_tree_num_nodes_leaves_max_depths(reg_tree)
tree_image = illustrate_trees(reg_tree, max_depth = 5, open_file = TRUE, length_in_px_per_half_split = 40) # will give the locally best nodes 

#in-sample stats
y_hat_train = predict(reg_tree, housing_Xtrain)
e = housing_ytrain - y_hat_train
sd(e) #s_e
insamp_r_sq = 1 - sd(e) / sd(housing_ytrain) #R^2
cat("in sample r^2 = ", insamp_r_sq , "\n ")

#oos stats
y_hat_test_tree = predict(reg_tree, housing_Xtest)
e = housing_ytest - y_hat_test_tree
sd(e)
oos_r_sq = 1 - sd(e) / sd(housing_ytest)
cat("oos r^2 =" , oos_r_sq )
```

```{r}
#Linear Modeling
pacman::p_load(xtable)

lin_mod = lm(housing_ytrain ~ . , housing_Xtrain)
lin_mod

#in-sample stats
summary(lin_mod)$sigma
lin_insample_rsq =summary(lin_mod)$r.squared
cat("insaple r^2 =" ,lin_insample_rsq, "\n")
xtable(lin_mod)

#oos stats
y_hat_test_linear = predict(lin_mod, housing_Xtest)
e = housing_ytest - y_hat_test_linear
sd(e)
lin_oos_rsq = 1 - sd(e) / sd(housing_ytest)
cat("oos r^2 =" , lin_oos_rsq, "\n")
summary(lin_mod)
```

```{r}
#Random Forest 


pacman::p_load(mlr)
#housing_Xcomplete = union_all(housing_Xtrain, housing_Xtest) # its iliegal to use the test data
housing_Xcomplete = housing_Xtrain
#y_salesprice = union_all(housing_ytrain, housing_ytest)
y_salesprice = housing_ytrain
mlr_data = cbind(y_salesprice, housing_Xcomplete)
colnames(mlr_data)[1] = "sales_price"
task = makeRegrTask(data = mlr_data, target = "sales_price")

parms = makeParamSet(
  makeIntegerParam("mtry", lower = 2, upper = ncol(housing_Xcomplete)), #feature depandent mtry can not be greater than num of col
  makeIntegerParam("ntree", lower = 2, upper = 100 ), # it is possible to go higher 
  makeIntegerParam("nodesize", lower =2, upper = 100)
)


desc <- makeResampleDesc("CV", iters = 20)
ctrl <- makeTuneControlRandom(maxit = 20)
mlr_ret <- tuneParams("regr.randomForest", task = task, resampling = desc, par.set = parms, control = ctrl, measures = list(rmse))

#Optimal result
mlr_ret

```

```{r}
rf_mod = YARF(housing_Xcomplete, y_salesprice, mtry= as.integer(mlr_ret$x[1]), num_trees = as.integer(mlr_ret$x[2]), nodesize =as.integer(mlr_ret$x[3]) )
rf_mod
#rf_is_mod = YARF(housing_Xtest, housing_ytest, mtry= as.integer(mlr_ret$x[1]), num_trees = as.integer(mlr_ret$x[2]))
rf_is_mod
yhat = predict(rf_is_mod, housing_Xtest)
```


```{r}
#gfinal oos for the final model 
cat("gfinal")
oos_rmse = sqrt(mean((housing_ytest - yhat)^2))
oos_rsq = 1 - sum((housing_ytest - yhat)^2)/sum((housing_ytest - mean(y_salesprice))^2)
oos_rmse
oos_rsq

```



